---
title: "model"
author: "benzi"
date: '2022-12-10'
output:
  pdf_document: default
  html_document: default
---

```{r}
#setwd("~/Heart disease analysis")
library("tidyverse")
data<-read.csv("heart.csv")
head(data)
```


```{r}
summary(data)
```


```{r}
colnames(data)
```
## Data transformation
```{r}
data2<-data %>%
  mutate(sex = if_else(sex==1,"MALE","FEMALE"),
         fbs = if_else(fbs==1,">120","<=120"),
         exang = if_else(exang==1,"YES","NO"),
         cp = if_else(cp==1,"ATYPICAL ANGINA",
                      if_else(cp==2,"NON-ANGINAL PAIN","ASYMPTOMATIC")),
         restecg = if_else(restecg==0,"NORMAL",
                           if_else(restecg==1,"ABNORMALITY","PROBABLE OR DEFINITE")),
         slope = as.factor(slope),
         ca = as.factor(ca),
         thal = as.factor(thal),
         target = if_else(target==1,"YES","NO")
         ) %>%
  mutate_if(is.character,as.factor) %>%
  dplyr::select(target,sex,fbs,exang,cp,restecg,slope,ca,thal,everything())
```

## Data visualization
# Barplot for target
```{r}
ggplot(data2,aes(x=target,fill=target))+
  geom_bar()+
  xlab("Heart disease")+
  ylab("count")+
  ggtitle("Presence & Abscence of Heart disese")+
  scale_fill_discrete(name="Heart disease", labels = c("Abscence","Presence"))
prop.table(table(data2$target))
```
# Count of the frequency of age
```{r}
data2%>%
  group_by(ï..age) %>%
  count() %>%
  filter(n>10) %>%
  ggplot()+
  geom_col(aes(ï..age,n),fill="green")+
  ggtitle("Age analysis")+
  xlab("Age")+
  ylab("Age count")
```
# Compare blood pressure across chest pain
```{r}
data2 %>%
  ggplot(aes(x=sex,y=trestbps))+
  geom_boxplot(fill = "Purple")+
  xlab("sex")+
  ylab("BP")+
  facet_grid(~cp)
```
```{r}
data2 %>%
  ggplot(aes(x=sex,y=chol))+
  geom_boxplot(fill = "orange")+
  xlab("sex")+
  ylab("BP")+
  facet_grid(~cp)
```
## Correlation
```{r}
library(corrplot)
library(ggplot2)
cor_heart<-cor(data2[,10:14])
cor_heart
corrplot(cor_heart,method = "square",type = "upper")
```
## Check for null values
```{r}
s = sum(is.na(data2));s
is.null(data2)
```
## Training and Testing data
```{r}
library(caret)
set.seed(10)
colnames(data2)
inTrainRows <- createDataPartition(data2$target,
                                   p=0.7,list=FALSE)
trainData <- data2[inTrainRows,]
testdata <- data2[-inTrainRows,]
nrow(trainData)/(nrow(testdata)+nrow(trainData))
# create 2 lists
AUC = list()
Accuracy = list()
```
## Logistic Regression
```{r}
set.seed(10)
logRegModel <- train(target ~ .,
                     data=trainData,
                     method = 'glm',
                     family = 'binomial')
logRegPrediction <- predict(logRegModel, testdata)
logRegPredictionProb <- predict(logRegModel,
                                testdata,
                                type='prob')[2]
logregConfMat <- confusionMatrix(logRegPrediction,
                                 testdata[,"target"])
# RDC curve
library(pROC)
AUC$logReg <- roc(as.numeric(testdata$target),
                  as.numeric(as.matrix(logRegPredictionProb)))$auc
# find names with str(logRegConfMat)
Accuracy$logReg <- logregConfMat$overall['Accuracy']
```
## Support Vector Machine
```{r}
fitControl <- trainControl(method = "repeatedcv",
                           number = 10,
                           repeats = 10,
                           classProbs = TRUE,
                           summaryFunction = twoClassSummary)
set.seed(10)
svmModel <- train(target ~ .,
                  data = trainData,
                  method = "svmRadial",
                  trControl = fitControl,
                  preProcess = c("center", "scale"),
                  tuneLength = 8,
                  metric = "ROC")
svmPrediction <- predict(svmModel, testdata)
svmPredictionProb <- predict(svmModel,
                             testdata,
                             type='prob')[2]
svmConfMat <- confusionMatrix(svmPrediction,
                              testdata[,"target"])
# RDC curve 
AUC$svm <- roc(as.numeric(testdata$target),
               as.numeric(as.numeric(as.matrix(svmPredictionProb))))$auc
Accuracy$svm <- svmConfMat$overall['Accuracy']
```
## Random Forest
```{r}
library(randomForest)
set.seed(10)
RFModel <- randomForest(target ~ .,
                        data = trainData,
                        importance = TRUE,
                        ntree = 200)
# varImpPlot(RFModel)
RFPrediction <- predict(RFModel,
                        testdata)
RFPredictionProb <- predict(RFModel,
                            testdata,
                            type = "prob")[,2]
RFConfMat <- confusionMatrix(RFPrediction,
                             testdata[,"target"])
# RDC curve 
AUC$RF <- roc(as.numeric(testdata$target),
               as.numeric(as.numeric(as.matrix(RFPredictionProb))))$auc
Accuracy$RF <- RFConfMat$overall['Accuracy']
```
## Comparison of AUC and Accuracy
```{r}
row.names <- names(Accuracy)
col.names <- c("AUC","Accuracy")
cbind(as.data.frame(matrix(c(AUC,Accuracy),nrow = 3,ncol = 2,
                           dimnames = list(row.names,col.names))))
```
```{r}
summary(logRegModel)$coeff
```
## Confusion Matrix
```{r}
logregConfMat
```
# the p-value is small, reject the proposed hypothesis, accept the alternative hypothesis
```{r}
svmConfMat
```
```{r}
RFConfMat
```

